Stack Safety as a Security Property

### 2020-07-08 main.tex
CP - 6. keeping lock

\definition
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}

\definition The testing property keeps track of relevant information
on a {\em variant stack}: a list of machine states, each element of
which (denoted $\vm{C}$) of which is a variation of the actual machine
according to a contour $C$.
%
\footnote{The full formal definition in Coq also records additional
  information (the initial machine state at the call, its initial
  variation, the contour, and a return predicate) which we elide
  here for simplicity of exposition.}
%%
%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
%  $$\vs \in \VSS$$
%%
\iflater
\leo{This can be elided}
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}
  \fi

@@

\definition{Eager testing property}
  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
  testing property} at a given call depth as follows.
% Refactor repetitive notation before case split
  In all cases that assume a machine step, $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, assume also (unless the step is a return) that the
  variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
  machines can step.

@@

@@

  \item
%
    If (1) the machine takes a \emph{call step} and $\contour$ is the contour at
    the call,
%
    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
    satisfied,
%
    (3) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

>>>>

\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
\emph{stack confidentiality test} if the machines they stepped to
have their changed components coincide and the observations during
these steps are equal:
\[
\begin{array}{rl}
  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
  \mach_1 \stepstoobs{\obs_1} \mach_1' \wedge \mach_2 \stepstoobs{\obs_2} \mach_2' \Rightarrow \\
&    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
\end{array}
  \]

\definition The testing property keeps track of a variation machine
for every nested call performed so far on a {\em variant stack}: a
list of machine states annotated with a contour, each element of which
(denoted $\vm{C}$) of which is a variation of the actual machine
according to a contour $C$.
%
\footnote{The full formal definition in Coq also records additional
  information (the initial machine state at the call, its initial
  variation, the contour, and a return predicate) which we elide
  here for simplicity of exposition.}
%%
%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
%  $$\vs \in \VSS$$
%%
\iflater
\leo{This can be elided}
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}
  \fi

@@

\definition
%  Let $\callmap$ a call map and $\codemap$ a function map.
  Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
  variant stack $\vs$, we say that an MP-trace induced from $(\mach,
  \pol)$ satisfies the \emph{strong stack safety} at a given call
  depth as follows.
% Refactor repetitive notation before case split
  In all cases that let $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$.
  %and assume (unless the step is a return) that the
  %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
  %machines can step.

@@

  \item
%
    If the machine takes a \emph{call step} and $\contour$ is the contour at
    the call then,
%
    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
    satisfied,
%
    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

@@
